---
title: "BDA - Project"
author: "Anonymous"
output:
  pdf_document:
    toc: yes
    toc_depth: 1
  word_document:
    toc: yes
    toc_depth: '1'
urlcolor: blue
---

```{r setup, include=FALSE, warning=FALSE}
# This chunk sets echo = TRUE as default, that is print all code.
# knitr::opts_chunk$set can be used to set other notebook generation options, too.
# include=FALSE inside curly brackets makes this block not be included in the pdf.
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

## The motivation
Annually many neonatal in the world dead because of different kind of reasons in this research we wanted to know what is the rate of neonatal death in Finland and Colombia countries in 2020 for different cities and states.




## The Problem
The problem was finding the neonatal probability of each Colombia state. 
because we treated number of total births as a number of cases and number of deaths as a successes. So we wanted to compare the death rate for different cities and satates.


## The main modeling idea 
The main modeling idea was separate model.



##  Some illustrative figure

```{r}
# Read Excel file and convert it to DataFrame
library("readxl")
library("rstan")

path <- file.path(getwd(),"Bayesian","Project", "NeonatalDeathColombia.xlsx" )
exceldata = read_excel("NeonatalDeathColombia.xlsx")                                                                            
dfData= data.frame(exceldata)
print(head(dfData))
deaths = dfData$Late.perinatal.and.neonatal.mortality.2020
births = dfData$Late.perinatal.and.neonatal.mortality.2020 + dfData$Live.births.to.2020
plot(deaths ~ births, data = dfData)
lm1 <- lm(deaths ~ births, data = dfData)
summary(lm1)
abline(lm1, col = 2, lty = 4, lw = 3)
path <- file.path(getwd(),"Bayesian","Project", "a.xlsx" )
exceldata = read_excel("a.xlsx")                                                                            
dfA= data.frame(exceldata)

path <- file.path(getwd(),"Bayesian","Project", "b.xlsx" )
exceldata = read_excel("b.xlsx")                                                                            
dfB= data.frame(exceldata)
```

# Description of the data and the analysis problem. Provide information where the data was obtained, and if it has been previously used in some online case study and how your analysis differs from the existing analyses.

We found data from Colombia government data website about neonatal mortality. https://www.ins.gov.co/Paginas/mapa-del-sitio.aspx. But we didn't have data in one excel file we created data and fill the birth and death information for each year finding them separatly. We didn't find any online analysis same as this before, so we analyzed this data from our point of view. We had different columns as number of live births and number of dead birth we added these two columns to find the number of total births.  


# Description of at least two models, for example:non hierarchical and hierarchical

We used a hierarchical and separate models for this data. We calculated the separate mean and sd for previous years. We initially used beta-binomial separate model and then improve that with beta-binomial hierarchical model.


# Informative or weakly informative priors, and justification of their choices.

# Stan, rstanarm or brms code.

```{r}
library(rstan)
library(bayesplot)
stan_data <- list(
  N = 37,
  L = 1,
  y = deaths,
  n = births,
  aMean = dfA$Mean,
  aStd = dfA$Std,
  bMean = dfB$Mean,
  bStd = dfB$Std
    
)
```

```{r}
write("// Stan Separate model
data {
  int<lower=0> N; // Number of states
  real L;
  int<lower=0> y[N]; //Number of neonatal deaths
  int<lower=0> n[N]; //Number of births
  real aMean[N]; //Minimun a value
  real aStd[N]; //Maximum a value
  real bMean[N]; //Minimun b Value
  real bStd[N]; //Maximum b value
}

// The parameters accepted by the model
parameters {
  vector<lower=0, upper=1>[N] p;
  vector<lower=L>[N] a;
  vector<lower=L>[N] b;
  
}

// The model to be estimated. We model the output
// 'y' to be normally distributed with mean 'mu'
// and standard deviation 'sigma'.
model {
  //Priors  
    for (i in 1:N) {
      a[i] ~ normal(aMean[i], aStd[i]) T[L,]; //Number of successes parameter
      b[i] ~ normal(bMean[i], bStd[i]) T[L,]; //Number of no Success
    }
    
    for (j in 1:N) {
      p[j] ~ beta(a[j], b[j]);
    }
      
    //Likelihood 
    for (k in 1:N){
      y[k] ~ binomial(n[k], p[k]);
    }

}

generated quantities {
    //Prediction capital city
    real ypred;
    //Log Likelihood ratios
    vector[N] log_lik;
    //Predictive distribution of the capital city
    ypred = binomial_rng(n[6],p[6]);
    
      for(j in 1:N){
        log_lik[j] = binomial_lpmf(y[j] | n[j], p[j]);
      }
      
}// The posterior predictive distribution",

"separate_model.stan")
```


```{r, include=FALSE}
stanc("separate_model.stan")
library(rstanarm)
sm <- rstan::stan_model(file = "separate_model.stan")
stan_model1 <- "separate_model.stan"

separate_model <- rstan::sampling(sm, data = stan_data)
monitor(separate_model)

separate_extract_log_lik <- extract_log_lik(separate_model, parameter_name = "log_lik", merge_chains = FALSE);
r_eff <- relative_eff(exp(separate_extract_log_lik), cores = 2) 
separate_model_loo <- loo(separate_extract_log_lik, r_eff = r_eff, cores = 2)
print(separate_model_loo)
hist(separate_model_loo$diagnostics$pareto_k, main="k values for separate model")


```


```{r}
write("// Stan Beta-binomial Hierarchical model 
data {
  int<lower=0> N; // Number of states
  real L;
  int<lower=0> y[N]; //Number of neonatal deaths
  int<lower=0> n[N]; //Number of births
  real aMean; //Mean a value
  real bMean; //Mean b Value
  real logn;

  
}

// The parameters accepted by the model
parameters {
  vector<lower=0, upper=1>[N] p;
  real mu;
  real loge;
}

transformed parameters{
  real a;
  real b;
  real e;
  
  e = exp(loge);
  a = mu*e;
  b = (1-mu)*e;
  
  
}

// The model to be estimated. We model the output
// 'y' to be normally distributed with mean 'mu'
// and standard deviation 'sigma'.
model {
  
  
  mu ~ beta(aMean,bMean);
  loge ~ logistic(logn, 1);
 
  
  //Prior
    for (j in 1:N) {
      p[j] ~ beta(a, b);
      
    }
  
    
    //Likelihood 
    for (k in 1:N){
      y[k] ~ binomial(n[k], p[k]);
    }

}

generated quantities {
    //Prediction capital city
    real ypred;
    //Log Likelihood ratios
    vector[N] log_lik;
    //Predictive distribution of the capital city
    ypred = binomial_rng(n[6],p[6]);
    
      for(j in 1:N){
        log_lik[j] = binomial_lpmf(y[j] | n[j], p[j]);
      }
      
    }// The posterior predictive distribution",

"hierachichal_model.stan")
```


```{r, include=FALSE}
library(rstan)
library(bayesplot)
s = 12345
sm_hierachichal <- rstan::stan_model(file = "hierachichal_model.stan")
l = log(100)
stan_data_hierachichal <- list(
  N = 37,
  L = 1,
  y = deaths,
  n = births,
  aMean = mean(dfA$Mean),
  bMean = mean(dfB$Mean),
  logn = l

)
stanc("hierachichal_model.stan")

sm <- rstan::stan_model(file = "hierachichal_model.stan")
stan_model1 <- "hierachichal_model.stan"

hierachichal_model <- rstan::sampling(sm_hierachichal, data = stan_data_hierachichal)
monitor(hierachichal_model)

hierachichal_extract_log_lik <- extract_log_lik(hierachichal_model, parameter_name = "log_lik", merge_chains = FALSE);
r_eff <- relative_eff(exp(hierachichal_extract_log_lik), cores = 2) 
hierachichal_model_loo <- loo(hierachichal_extract_log_lik, r_eff = r_eff, cores = 2)
print(hierachichal_model_loo)
hist(hierachichal_model_loo$diagnostics$pareto_k, main="k values for hierachichal model")

posterior <- extract(hierachichal_model)

```

```{r}
plot(deaths ~ births, data = dfData)




abline(mean(posterior$a), mean(posterior$b), col = 3, lw = 2)
abline(mean(posterior$a), mean(posterior$b), col = 36, lw = 3)
```

```{r}
plot(posterior$alpha, type = "l")
plot(posterior$beta, type = "l")
plot(posterior$sigma, type = "l")
```

# How to the Stan model was run, that is, what options were used. This is also more clear as combination of textual explanation and the actual code line.

# Convergence diagnostics (Rˆ, ESS, divergences) and what was done if the convergence was not good with the first try.


# Posterior predictive checks and what was done to improve the model.

# Model comparison (e.g. with LOO-CV).

# Predictive performance assessment if applicable (e.g. classification accuracy) and evaluation of practical usefulness of the accuracy. If not applicable, then explanation why in this case the predictive performance is not applicable.



# Sensitivity analysis with respect to prior choices (i.e. checking whether the result changes a lot if prior is changed)


# Discussion of issues and potential improvements.


# Conclusion what was learned from the data analysis.



# Self-reflection of what the group learned while making the project.

